Generative Adversarial Networks

### 2017-11-28 cs6735/project/draft.tex
continued report
updated cost

Generative adversarial networks (GANs), first introduced in 2014, offer a new framework for estimating generative models with use of an adversarial process \citep{2014arXiv1406.2661G}. These models offer a clever approach to solving the aforementioned problem of explicitly defining a cost function. Rather than training a single model, a discriminator is introduced. These two networks are pitted against each other in a minimax game. The generative model $G$ attempts to produce data that resembles that of the training set while the discriminative model $D$ attempts to classify whether or not a sample was real of generated. To train this network, $G$ attempts to maximize the error rate of $D$ whereas $D$ attempts to minimize it. Backpropagation is used to update the parameters and to train each model. In the ideal case, $D$ is unable to distinguish the generated samples from the real samples and produces an error rate of $\frac{1}{2}$ \citep{2014arXiv1406.2661G}.

@@

\begin{figure}
\centering
\includegraphics[scale=0.65]{img2img}
\caption{Examples of the image-to-image technique developed by Isola \textit{et al.} \citep{2016arXiv161107004I}. This approach makes use of conditional generative adversarial networks to map image domains.}
\label{fig:Image to Image Examples}
\end{figure}

@@

$$\min_{G}\max_{D}V(D, G) = \mathbb{E}_{x\sim p_{data}(x)}[\log(D(x))] + \mathbb{E}_{z\sim p_z(z)}[\log(1-D(G(z)))]$$

@@

$$J_d(\theta_D,\theta\G) = -\frac{1}{2}\mathbb{E}_{x\sim p_{data}(x)}[\log(D(x))] - \frac{1}{2}\mathbb{E}_{z\sim p_{z}(z)}[\log(1-D(G(z)))]$$

This is a standard cross entropy cost function that is trained on two minibatches, one from the dataset and one from the generator. For this to become a zero-sum game, a minimax game, the cost function of the generator is simply the negation of the discriminator.

$$J_g(\theta_D,\theta\G) = -J_d(\theta_D,\theta\G)$$

This version of the generator loss function is particularly useful for theoretical analysis as shown in the first GAN paper \citep{2014arXiv1406.2661G}. It can be shown that the minimax game is the same as minimizing the Jensen-Shannon divergence and that each player will eventually converge if both players are able to directly update the function space as oppose to the parameters within a neural network \citep{2017arXiv170100160G}. However, in this minimax version, the loss function of the generator suffers from a vanishing gradient when the discriminator confidently rejects the generated data. The solution proposed by Goodfellow is to train the generator to maximize $log(D(G(z)))$ rather than minimize $log(1-D(G(z)))$ \citep{2014arXiv1406.2661G}. This function is heuristically motivated as to provide strong gradients when either the generator or discriminator is not functioning optimally.

$$J_g(\theta_D,\theta\G) = -\frac{1}{2}\mathbb{E}_{z\sim p_{z}(z)}[\log(D(G(z)))]$$

Maximum likelihood learning is another option for training GANs \citep{2017arXiv170100160G}. In this situation, we would be minimizing the Kullback-Leibler divergence between $p_{data}$ and $p_{model}$.

\begin{figure}
\centering
\includegraphics[scale=0.65]{costfunc}
\caption{The cost of the generator................}
\label{fig:Image to Image Examples}
\end{figure}

@@

When this occurs, the generator attempts to map every latent vector to the one $x$ variable the discriminator is most likely to guess is real. When training, one can only hope the behavior resembles the minimax version. It is often speculated that the root cause of the issue the particular divergence measure used. Specifically, it is often asserted that mode collapse is caused by the use of Jenson-Shannon divergence; however, the use of Kullback-Leibler divergence has be shown to worsen the problem \citep{2017arXiv170100160G}. As mode collapse is currently an unavoidable reality for GANs, their applications are limited to situations where it is acceptable to generate only a small number of distant samples \citep{2017arXiv170100160G}. A potential solution to this problem is discussed in the following section.

@@

\subsection{Discrete Data}

>>>>

Generative adversarial networks (GANs), first introduced in 2014, offer a new framework for estimating generative models with use of an adversarial process \citep{2014arXiv1406.2661G}. These models offer a clever approach to solving the aforementioned problem of explicitly defining a cost function. Rather than training a single model, a discriminator is introduced. These two networks are pitted against each other in a minimax game. The generative model $G$ attempts to produce data that resembles that of the training set while the discriminative model $D$ attempts to classify whether or not a sample was real of generated. To train this network, $G$ attempts to maximize the error rate of $D$ whereas $D$ attempts to minimize it. The goal of the GAN algorithm is to reduce the distance between the data and generated probability distributions. Backpropagation is used to update the parameters and to train each model. In the ideal case, $D$ is unable to distinguish the generated samples from the real samples and produces an error rate of $\frac{1}{2}$ \citep{2014arXiv1406.2661G}.

@@

\begin{figure}
\centering
\includegraphics[scale=0.55]{img2img}
\caption{Examples of the image-to-image technique developed by Isola \textit{et al.} \citep{2016arXiv161107004I}. This technique makes use of conditional generative adversarial networks to map across image domains.}
\label{fig:Image to Image Examples}
\end{figure}

@@

$$\min_{G}\max_{D}V(D, G) = \mathbb{E}_{x\sim p_{data}(x)}\Big[\log(D(x))\Big] + \mathbb{E}_{z\sim p_z(z)}\Big[\log(1-D(G(z)))\Big]$$

@@

$$J^{(D)}(\theta_D,\theta\G) = -\frac{1}{2}\mathbb{E}_{x\sim p_{data}(x)}\Big[\log(D(x))\Big] - \frac{1}{2}\mathbb{E}_{z\sim p_{z}(z)}\Big[\log(1-D(G(z)))\Big]$$

This is a standard cross entropy cost function that is trained on two minibatches, one from the dataset and one from the generator. For this to become a zero-sum game, the cost function of the generator is simply the negation of the discriminator.

$$J^{(G)}(\theta_D,\theta\G) = -J_d(\theta_D,\theta\G)$$

This version of the generator loss function is particularly useful for theoretical analysis as shown in the first GAN paper \citep{2014arXiv1406.2661G}. Goodfellow demonstrated that the minimax game is the same as minimizing the Jensen-Shannon divergence and that each player will eventually converge if both players are able to directly update the function space as oppose to the parameters within the network \citep{2017arXiv170100160G}. However, in this minimax version, the loss function of the generator suffers from a vanishing gradient when the discriminator confidently rejects the generated data. The solution proposed by Goodfellow is to train the generator to maximize $log(D(G(z)))$ rather than minimize $log(1-D(G(z)))$ \citep{2014arXiv1406.2661G}. This function is heuristically motivated as to provide strong gradients when either the generator or discriminator is not functioning optimally. In this version, the generator attempts to maximize the log probability of a misclassification whereas the minimax version attempts to minimize the log probability of the discriminator being correct.

$$J^{(G)}(\theta_D,\theta\G) = -\frac{1}{2}\mathbb{E}_{z\sim p_{z}(z)}\Big[\log(D(G(z)))\Big]$$

A maximum likelihood cost is another option for training GANs \citep{2017arXiv170100160G}. It can be shown that that this cost function minimizes the Kullback-Leibler divergence between $p_{data}$ and $p_{model}$ instead of the Jensen-Shannon divergence. This cost estimator is useful as it allows for comparison with other generative models; however, it behaves similarly to the minimax cost function as seen in \ref{fig:CostFunc}.

$$J^{(G)}(\theta_D,\theta\G) = -\frac{1}{2}\mathbb{E}_{z\sim p_{z}(z)}\Big[\exp(\sigma^{-1}(D(G(z))))\Big]$$

\begin{figure}
\centering
\includegraphics[scale=0.5]{costfunc}
\caption{A comparison of the generator cost function for generative adversarial networks. When $D(G(z))$ is close to $0$, the gradient is saturated for both the minimax and maximum likelihood cost functions. This situation occurs when the discriminator $D$ confidently rejects the generated value $G(z)$ as fake. Therefore, the non-saturated heuristic cost function is the only one which works well in practice. Image retrieved from Goodfellow NIPs tutorial \citep{2017arXiv170100160G}.}
\label{fig:CostFunc}
\end{figure}

@@

When this occurs, the generator attempts to map every latent vector to the one $x$ variable the discriminator is most likely to guess is real. When training, one can only hope the behavior resembles the minimax version. It is often speculated that the root cause of the issue the particular divergence measure used. Specifically, it is often asserted that mode collapse is caused by the use of Jensen-Shannon divergence; however, the use of Kullback-Leibler divergence has be shown to worsen the problem \citep{2017arXiv170100160G}. As mode collapse is currently an unavoidable reality for GANs, their applications are limited to situations where it is acceptable to generate only a small number of distant samples \citep{2017arXiv170100160G}. A potential solution to this problem is discussed in the following section.

@@

\subsection{Discrete Data}
GANs assume a model implementation that is differentiable due to it's use of back propagation. Therefore, this restricts their ability to generate discrete data.
